---
title: "pubh7462_hw1_RMD"
author: Tsz Fung Wong
date: January 28,2022
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
#Load the good stuff
library(tidyverse)
library(ggplot2)
#Working directory for .RMD
knitr::opts_knit$set(echo = TRUE,
root.dir = rprojroot::find_rstudio_root_file())
#Controlling figure output in markdown
knitr::opts_chunk$set(
# fig.height = 4
fig.width = 6,
# fig.asp = .5,
out.width = "90%",
# out.height =
fig.align = "center",
cache = FALSE,
echo = TRUE
)
#Set Theme for ggplot2
theme_set(theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom"))
#Set Scientific notation output for knitr
options(scipen = 999)
```

# Question 1

```{r}
q1data = tibble(x = rnorm(1000,0,1), y = rnorm(1000, 1, 2))
q1data = q1data %>% 
  mutate(sum_indicator = ifelse(q1data$x + q1data$y > 0.5, TRUE, FALSE)) %>%
  mutate(indicatorTRUE = forcats::fct_recode(factor(sum_indicator),"Yes" = "TRUE" ,"No" = "FALSE"))
q1data$indicatorTRUE = forcats::fct_relevel(q1data$indicatorTRUE, "Yes")
ggplot(q1data, aes(x, y, color = factor(indicatorTRUE))) + geom_point() +
  labs(title="Random Normal Samples", x ="X ~ N(0, 1)", y ="Y ~ N(1, 2)") +
  scale_color_discrete("X + Y > 0.5", q1data$indicatorTRUE)
```

# Question 2

## Description of Penguins Dataset
There are 344 observations of penguines, each is described using 8 variables. The dataset contains some missing entries denoted by NA.  

## Variables Meaning  
The first column contains species information for the penguins classified the penguines into three species, namely Adélie, Chinstrap and Gentoo. The second column contains geographic information about the penguins, it describe the island in Palmer Archipelago, Antarctica that the penguines belongs to (Biscoe, Dream or Torgersen). Followed by biological statistics about the penguins, bill length, bill depth, flipper length are ordinal variables in milimeters, body mass is another ordinal variable in grams. Sex for the penguines is also included and is classified into female and male. The final variable is the study year record, takes value either in 2007, 2008 or 2009. 

```{r}
penguin.df <- read_rds("./data/penguin.RDS")
flipper_bill_length = ggplot(na.omit(penguin.df), aes(bill_length_mm, flipper_length_mm, color = species)) + geom_point() + 
  labs(title="Bill length verses Flipper length", x ="Flipper length(mm)", y ="Bill length(mm)") +
  scale_color_discrete("Species",penguin.df$species)
flipper_bill_length
flipper_bill_length + facet_wrap(~ sex)

```

## Observation from the plots

It seems that throughout the three species of the penguins, all shows positive correlation between length of flipper and bill. However, when the plots are separated by  sex, there is no evident trend that the length of flipper and bill are correlated. The seemingly possible trend in the first plot is owing to the gender difference between male and female penguins, which male tends to have a longer flipper and bill compared to female ones of the same species.  